### 2013-07-06 bibliography.tex

% This is bibliography.tex

% If you don't have a bibliography, use the "\endinput" command here.
% \endinput

\begin{thebibliography}{99}

% Write your bibliography here using the \bibitem{} command. 
% Be sure to strictly follow your bibliography formatting system (eg. Vancouver, Harvard, Oxford, or IEEE-system).

%Example in shown below:

\bibitem{digital imaged based tree measurement for forest inventory}
Juujärvi J, Heikkonen J, Brandt S, Lampinen J. Digital Image Based Tree Measurement for Forest Inventory. Proceedings of SPIE. 1998; Volume 3522: 114-123.

\end{thebibliography}