### 2013-07-06 paper.tex

%This is paper.tex

%Write the contents of your paper here.

\newcommand{\degree}{$^\circ$}

\section{Introduction}
Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory. The use of this information is important to companies who need to make smart, profitable decisions but also need to be certain that the choices they make now are maintable in the future. However this information also benefits climate modellers, environmental groups and governments who want to be able to sustain ecosystems and plan restorations. To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis. Most methods used today to measure individual trees are often done manually with a variety of tools. The use of these analogue approaches is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}. SilviaTerra, however, is a company willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results. This is achieved through computer exactitude combined with statistical analysis. Recently a new project was started, with this research paper developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.

The measurements required for forest inventory can be achieved through a number of methods. One of them is using a so called biltmore stick. A biltmore stick is similar to a normal ruler but has been specially modified for measuring trees. When standing at a certain distance the user can discern the height of the tree by holding the biltmore stick up to their eye and reading the corresponding number. The biltmore stick can also easily be used to measure the DBH of the tree. Other ways of measuring tree volume include using a combination of lasers and inclinometers or by having a person manually climb up the tree and drop a tape measure down to measure the height. Despite these relatively simpe ways of measuring they often leave room for error. That’s why a more common way of getting a precise reading is to cut down trees in the area and then physically measure them. After the relevant data is collected one applies a stem-profile taper equation that is specific to the species of tree measured. These taper functions are based on numerous measurements and have been refined over the years. The main problem with taper functions is that to get an accurate result they need to be callibrated to the hundreds of local conditions. This again is cost prohibitive and only so much of a forest can be accounted for in the taper function. Even when one manages to create a new taper function, it is still far from perfect as all trees, even from the same species, differ in shape.

This project aims to solve the described problem by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation. This is supposed to then give a accurate result, cheaply and non-destructively. All within the reaches of a mobile phone.

\section{Background}

\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.65\textwidth]{main.png}
	  	\caption{Homescreen}
  		\label{main}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.65\textwidth]{input.png}
	  	\caption{User input}
  		\label{input}
	\endminipage\hfill
\end{figure}
The first thing a person does when accessing the mobile application is to log into the system with their user name and password. After that they are prompted with a simple menu where they can either choose to add new trees to their inventory or upload them to SilviaTerra's cloud service, figure \ref{main}. The data can then later be used with SilviaTerra's other products. If the user chooses to add a new tree then they will be asked to enter the tree's DBH and species, \ref{input}. 
\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.75\textwidth]{dbh.png}
	  	\caption{DBH picture}
	  	\label{dbh}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.75\textwidth]{bole.png}
	  	\caption{Bole picture}
	  	\label{bole}
	\endminipage\hfill
\end{figure}

After that they take two pictures. The interface is quite simple with a cross-hair to show where to point the camera. The first picture is of the actual DBH, figure \ref{dbh}, and the second picture is of the bole of the tree, figure \ref{bole}. Between these two pictures the camera will have to be tilted vertically to capture the bole. The angle between these two camera positions is recorded through the internal gyroscope. After this stage the user's role in the application is done and the pictures is sent to one of SilviaTerra's web servers. Here horizontal lines that run parallel with the bottom of the image are drawn across the entire picture. These are drawn at different pixel heights with equal pixel distance in between. The reason this is done is because computer vision is still an experimental technology and is not fully capable yet of being to detect the edges of trees with a background of forestry. Seeing as this is a commercial product, it would be quite disadvantageous if errors were commonplace. So instead the images with the newly drawn lines are then forwarded to Amazon's Mechanical Turk service. Just like the name suggests Amazon's Mechanical Turk is meant to be analogy to the 18th century concept of a automated chess board. The service is in its simplest form a crowd sourcing internet marketplace where individuals or businesses can request tasks to be completed. Usually the tasks are ones that are difficult for computers but easy for humans. SilviaTerra creates tasks where people manually click where the previously drawn lines, cross the edges of the tree. Ones of the points of intersection are identified, their corresponding pixel coordinates are the sent back to the web server. The pixel coordinates along with information from the phone is piped into a program where the volume is then calculated. It is this program that is the main topic of this research paper.

\section{Method}
Once all the required information has been retrieved it is piped into the program that will be calculating the volume. The dataset that is used comprises information about the phone and the results from Mechanical Turk. It is as follows:
\begin{itemize}
	\item DBH length
	\item Horizontal and vertical angle of view
	\item Horizontal and vertical image resolution
	\item Pitch of phone when images are captured
	\item Left and right pixel coordinates from Mechanical Turk
\end{itemize}

\subsection{Distance to tree}
The DBH is used as a reference to find all the other measurements. The first unknown that must be discovered is the distance to the tree. This distance is parallel with the ground. The difference between the x-coordinates of the left and right pixel that correspond to the DBH is calculated, figure \ref{horizontal_triangle}, $X_1$ and $X_2$. Due to the fact that the width of the DBH is now known in both pixels and in a physical measurement, the ratio between the two can be determined by dividing the physical distance between $X_1$ and $X_2$ with the number of pixels betwee them.
\begin{figure}[htp]
\centering
{
	\includegraphics[scale=0.3]{horizontal_triangle.pdf}
	\caption{Birds eye view of horizontal plane}
	\label{horizontal_triangle}
}
\end{figure}
 As seen in figure \ref{horizontal_triangle} an isosceles triangle can be constructed with the horizontal angle of view as the vertex angle. The triangles two congruent legs extend out towards the base of the triangle, $W_1$, which is inline with the tree. To get the length of the base, $W_1$, the previously calculated scale between pixels and the real world is multiplied with the number of pixels in the horizontal resolution of the image. The distance to the tree can easily calculated by drawing an extra line that bisects the vertex angle, contriving the angle $\theta$ which is exactly half of the angle of view and the actual distance to the tree $d$. This line also divides $W_1$ into $W_2$ and futhermore producing a right angle triangle that can be solved. Finally to actually calculate the distance, $d$, $W_2$ is divided by $\tan{\theta}$.

\subsection{Heights and widths} 
\begin{figure}[!htb]
	\minipage{0.45\textwidth}
		\centering
  		\includegraphics[width=0.9\textwidth]{triangle1.pdf}
	  	\caption{DBH picture}
	  	\label{triangle1}
	\endminipage\hfill
	\minipage{0.45\textwidth}
		\centering
	  	\includegraphics[width=0.9\textwidth]{triangle2.pdf}
	  	\caption{Bole picture}
	  	\label{triangle2}
	\endminipage\hfill
\end{figure}

The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale which is the ratio between degrees and pixels in the vertical axis is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane. With this method it is possible to find out the angle to each pixel coordinate, $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the second image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them. This is accomplished by employing the method for calculating the distance to the tree, in reverse. In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement. To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume. The current implemented method of doing this is by integrating all the points using the trapezoidal rule. All the widths are divided by 2 to get the radius. Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height. The basal area is plotted along with their corresponding heights. The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$. The area between all the points is the summed together.

This whole process is looped through for every identified pixel coordinate set until the width and height is known for all of them. Once that is done, a new loop is created for calculating the total volume.

\subsection{Simulation}
To test this method a \emph{virtual tree} was constructed. This tree was made out of blue tape and tapered off so that the program could be thoroughly tested. Along the tape points were marked out and the height to them plus the width between corresponding points were measured. Just like in the real application, two pictures were taken. One of the DBH and one of the top of the tree. Each set of points was then identified on the pictures and measured in pixels. The parameters were then passed into the program and the results were cross-referenced with the real life measurements.

\newpage

\section{Results}
\begin{table}[h!]
	\begin{center}
		\begin{tabular}{| l c c c c c c r |}		
		\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		\hline
		0 		& 0 	& 0 	& 0 	& 0 	& 0 	& 0 	& 0 	\\
		103 	& 96 	& 100 	& 103 	& 106 	& 113 	& 116 	& 123 	\\
		115 	& 109 	& 113 	& 116 	& 118 	& 125 	& 128 	& 135 	\\
		125 	& 125 	& 129	& 132 	& 134 	& 140 	& 144 	& 150 	\\
		159 	& 142 	& 147 	& 150 	& 151 	& 158 	& 161 	& 167 	\\
		166 	& 158 	& 163 	& 166 	& 167 	& 174 	& 175 	& 183 	\\
		183 	& 173 	& 179 	& 181	& 182 	& 189 	& 192 	& 198 	\\
		\hline
		\end{tabular}
		\caption{Heights over DBH}
		\label{heights}
	\end{center}
\end{table}

\begin{table}[h!]
	\begin{center}
		\begin{tabular}{| l c c c c c c r |}
		\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		\hline
		12       & 11.7      & 12.1      & 12.2      & 12.1      & 12.3      & 12.6      & 11.9      \\
		27       & 26.8      & 28.0      & 28.4      & 28.2      & 28.5      & 28.4      & 28.1      \\
		37       & 36.4      & 37.5      & 38.0      & 37.7      & 38.2      & 38.1      & 37.7      \\
		44       & 42.4      & 43.6      & 44.4      & 44.0      & 44.4      & 44.0      & 44.4      \\
		52       & 51.3      & 52.7      & 53.5      & 53.0      & 53.4      & 52.9      & 53.2      \\
		61       & 60.9      & 62.2      & 63.1      & 62.1      & 62.8      & 62.5      & 61.1      \\
		59       & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      & 59.0      \\
		\hline
		\end{tabular}
		\caption{Widths over DBH}
		\label{widths}
    \end{center}
\end{table}

\begin{table}[h!]
	\begin{center}
    	\begin{tabular}{| l c c c c c c r |}
    	\hline
		Measured & 57\degree & 47\degree & 39\degree & 27\degree & 22\degree & 18\degree & 15\degree \\
		412222   & 386642    & 396386    & 418370    & 412084    & 433294    & 445401    & 445624    \\
		\hline
		\end{tabular}
		\caption{Total volume over DBH}
		\label{volumes}
    \end{center}
\end{table}